DGE's and Enrichment

### 2022-01-31 reports/rmarkdown/manuscript/Figure4-5.Rmd
minor changes
change txt size and update igrap network

```{r Across tables, fig.height=4, fig.width=7, warning=F, include=FALSE}

TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
TRX_counts <- round(2^TRX - 1)
TRX_counts <- TRX_counts [ rowSums(TRX_counts>0)>= 1 , ]
TRX_counts <- TRX_counts[,sample_use]

@@

  image( t(top_TRX_counts[h$order,pull(df, "ID")][nrow(top_TRX_counts):1,]),
         col = colorRampPalette(c("navy","navy","grey95","firebrick4","firebrick4") )(99),
         breaks = seq(-5,5,length.out = 100),axes=F)
  mtext("           Differentialy Expressed Genes", side=3, las = 1, cex = .6, line =1.5,xpd=T, font = 2)
  title(xlab="Study Groups", line=1)

@@

  # GO
  res_list <- pluck(enrich, "GO", "res_list")
  #top_TRX_counts <- pluck(enrich, "GO", "top_TRX_counts")[,samples]

@@

    par(mar = c(1.2, 4.7, 2.2, 1))  #bottom, left, top, right
    barplot( c(rep(NA, each=20-length(comb)), comb),  
             xlim=c(-1.1,1.1),
             # xlim=c(min(c(comb,-2/abs(max(abs(DD))) ))*1.1, 
             #                          max(c(comb,2/abs(max(abs(DD)))) )*1.1), 
                        cex.names = .9, #mar=c(1,25,2,2), #main=paste(x, i, sep="\n"),
             names.arg = c(rep("", each=20-length(comb)),names_DD, names_UU ), 
             horiz = T,las= 1, border=NA,xaxt='n',line=-.2, #xlab="-log10 p-value",
             col = c(rep(NA, each=20-length(comb)), 
                     rev(colorRampPalette( c("firebrick3","gray","navy") )(19))[(comb+1)*9+1])
    )
    title(main=paste(x, str_replace_all(i, "_"," "), sep="\n"), line = 0.1, cex.main=.9)

    if(x == "L1-L2" | x == "T1-T2"){add_letter("b")}
    points_y <- rev(seq(par("usr")[4]-1.45, by=-1.2, length=length(c(UU,DD))))
             points( comb , points_y,
                     #mar=c(1,20,2,2),
                     bg=colorRampPalette( c("navy","gray","firebrick3") )(19)[(comb+1)*9+1],
                     pch=21, col="black",cex=1.5)
             lines(x=c(0, 0), y= c(points_y[1]-.6, par("usr")[4]-1)) 
             lines(x=c(-2/abs(max(abs(DD))),-2/abs(max(abs(DD)))), 
                   y= c(points_y[1]-.6, par("usr")[4]-1), lty=c(3)) # pval=0.01
             lines(x=c(2/abs(max(abs(UU))),2/abs(max(abs(UU)))), 
                   y= c(points_y[1]-.6, par("usr")[4]-1), lty=c(3))  # pval=0.01
             text(0,points_y[1]-1.5, "-log10 p-value", xpd=T)
    }
  }
}

dev.copy2pdf(file=paste0("./test.pdf"),
    width = 6.6,
    height = 6, paper = "a4"
    )
```

@@

sheets <- excel_sheets(paste0(out, DEG_files[1])) %>% set_names(.) %>% .[-c(1:2)]
arrange_plots.fun(enrich_L, "Luminal_gr", "Luminal", sheets, "TF_PPI_enrichment", enrichR_files[1], Layout = l)
```

@@

```{r UMAP, eval=FALSE, fig.height=2.5, fig.width=3, include=FALSE}
# , fig.height=3, fig.width=3.5
###############
#  TRX UMAP   #
###############
plot_UMAP.fun <- function(top_TRX_counts, group){
  #par(mar=c(4,4,2,6)) #b,l,t,r
  par(mar=c(2.5,2.5,2,3)) #b,l,t,r
  #par(mar=c(10,12,2,12)) #b,l,t,r
  set.seed(1)
  UMAP_TRX <- uwot::umap(t(top_TRX_counts),n_neighbors = 10,
                   metric = "correlation",min_dist = 0.1,spread = 5,
                   negative_sample_rate = 10)
  plot(UMAP_TRX,bg=pal[factor(group)],main="UMAP on top DGE RNAseq",
       pch=21,xlab="UMAP1",ylab="UMAP2",frame=F,axes=F)
  mtext("UMAP1", side=1, line=0.7, cex=0.8)
  mtext("UMAP2", side=2, line=0.7, cex=0.8)
  legend(par("usr")[2],par("usr")[4],title.adj = 0,
         #legend = levels(factor(metadata$Luminal_gr)),
         legend = sort(unique(group)),
         xjust = 0,yjust = 1, cex=0.7,
         bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)
}

plot_UMAP.fun(enrich_L$GO$top_TRX_counts, metadata$Luminal_gr)
m <- filter(metadata, !is.na(Tissue_gr)) 
plot_UMAP.fun(enrich_T$GO$top_TRX_counts[,m$ID], m$Tissue_gr)

#plot_UMAP.fun(enrich_T$GO$top_TRX_counts, metadata$Tissue_gr)
dev.copy2pdf(file=paste0("./UMAP",".pdf"),
    width = 4,
    height = 3.5, paper = "a4" #units = "cm", res = 300 
    )

@@

 <!-- -->

>>>>

```{r Across tables, fig.height=4, fig.width=7, warning=F, include=FALSE}
TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
TRX_counts <- round(2^TRX - 1)
TRX_counts <- TRX_counts [ rowSums(TRX_counts>0)>= 1 , ]
TRX_counts <- TRX_counts[,sample_use]

@@

  image( t(top_TRX_counts[h$order,pull(df, "ID")][nrow(top_TRX_counts):1,]),
         col = colorRampPalette(c("navy","navy","grey95","firebrick4","firebrick4") )(99),
         breaks = seq(-5,5,length.out = 100),axes=F)
  mtext("                                                          Differentially expressed genes and KEGG/GO pathways", side=3, las = 1, cex = .7, line =1.5,xpd=T, font = 2)
  title(xlab="Study Groups", line=1)

@@

  # GO
  res_list <- pluck(enrich, "GO", "res_list")

@@

    par(mar = c(1.2, 4.7, 2.2, 1))  #bottom, left, top, right
    barplot( c(rep(NA, each=20-length(comb)), comb),  
             xlim=c(-1.1,1.1), cex.names = .9,
             names.arg = c(rep("", each=20-length(comb)),names_DD, names_UU ), 
             horiz = T, las= 1, border=NA, xaxt='n', line=-.2,
             col = c(rep(NA, each=20-length(comb)), 
                     rev(colorRampPalette( c("firebrick3","gray","navy") )(19))[(comb+1)*9+1])
    )
    title(main=paste(x, str_replace_all(i, "_"," "), sep="\n"), line = 0.1, cex.main=1)

    if(x == "L1-L2" | x == "T1-T2"){add_letter("b")}
    points_y <- rev(seq(par("usr")[4]-1.45, by=-1.2, length=length(c(UU,DD))))
             points( comb , points_y,
                     bg=colorRampPalette( c("navy","gray","firebrick3") )(19)[(comb+1)*9+1],
                     pch=21, col="black",cex=1.5)
             lines(x=c(0, 0), y= c(points_y[1]-.6, par("usr")[4]-1)) 
             lines(x=c(-2/abs(max(abs(DD))),-2/abs(max(abs(DD)))), 
                   y= c(points_y[1]-.6, par("usr")[4]-1), lty=c(3)) # pval=0.01
             lines(x=c(2/abs(max(abs(UU))),2/abs(max(abs(UU)))), 
                   y= c(points_y[1]-.6, par("usr")[4]-1), lty=c(3)) # pval=0.01
             text(0,points_y[1]-1.5, "-log10 p-value", xpd=T)
    }
  }
}
```

@@

sheets <- excel_sheets(paste0(out, DEG_files[1])) %>% set_names(.) %>% .[-c(1:2)]
arrange_plots.fun(enrich_L, "Luminal_gr", "Luminal", sheets, "TF_PPI_enrichment", enrichR_files[1], Layout = l)

# test figure
# dev.copy2pdf(file=paste0("./test.pdf"),
#     width = 6.6,
#     height = 6, paper = "a4"
#     )
```

@@

```{r Suppl. Tbl. 12, include=FALSE}
metadata <- metadata %>% 
  mutate(L_T_groups = paste0(.$Luminal_gr, .$Tissue_gr)) %>%
  mutate(L_T_groups = ifelse(grepl("L2T2|L2T3", .$L_T_groups), .$L_T_groups, NA))

dge_LT <- edgR_dge.fun("L_T_groups", confound=T)
enrich_LT <- db %>%
  map(., ~gene_mod_enrich.fun(dge_LT$top,dge_LT$y,dge_LT$c,
                              db = .x,
                              group_name = "L_T_groups",
                              cuttoff = c(GO=2, KEGG=2) # c(GO=2, KEGG=2.7)
                              ))

# Plot the L2T2 vs L2T3 heatmap:
#pal <- c("#54ba63","#ad58c6","#d03e7a","#4cb6a9","#bb6fac","#cca442","#8fb53c","#6a7ecd","#63813e","#c46570","#b3743e","#d35238")
#plot_heatmap.fun(enrich_LT, "KEGG", "L_T_groups", "L_T_groups")

t <- dge_LT$top %>% filter(PValue <= 0.01)

df_across <- enframe(list(LtoT=enrich_LT)) %>%  
  unnest_longer(value) %>%
  mutate(res = map(.$value, "res_list")) %>%
  unnest_longer(res) %>%
  mutate(res = map(res, ~rownames_to_column(.x, var = "Terms"))) %>%
  mutate(res = map(res, ~filter(.x, pvalue <= 0.01))) %>%
  group_by(name, value_id) %>%
  mutate(mod = row_number()) %>% ungroup() %>%
  mutate(res = set_names(.$res, paste0(.$value_id,"_gene_module_",.$mod))) %>%
  group_split(name) %>% set_names(., map(., ~unique(.x$name)))

pages <- list(list("DEGs L2T2 vs L2T3"=t), df_across[[1]]$res) %>% flatten()
write.xlsx(pages, file=paste0(out,"Suppl.Tbl.12 L2T2_L2T3_DEGs_and_enrichment.xlsx")) 
```

## igraph network plot
```{r igraph network, echo=FALSE, warning=FALSE} 
# 1. Change the groups of interest by assigning condition_use
# 2. Change the number of genes to be included in the network by changing n
#    look at tempD_i and tempU_i to see how many genes was significant
# 3. Adjust the size of the gene nodes and text in the graph by changing; square, circle and txt

# combinations
sheets <- map(DEG_files, ~excel_sheets(paste0(out, .x))[-c(1:2)] ) %>%
  map(., ~set_names(.x))

# Set enrichment conditions
n <- 10  # maximum number of TF's to plot
# sizes:
square <- 10
cricle <- 8
txt <- .7

# Create a graph from the list of UP and DOWN regulated genes
i_graph <- function(temp_i, n, square, cricle){
  temp <- lapply( 1:n, function(x){ strsplit(as.character(temp_i$Genes[x]),split = ";")[[1]]  } )
  temp <- map(temp, purrr::discard, is.na) %>% compact() # removes NA
  names(temp) <- temp_i$`Transcription Factor`[1:length(temp)]
  temp2 <- lapply( names(temp) , function(x){ rep(x,length(temp[[x]])) })
  temp <- cbind(unlist(temp2),unlist(temp))
  g <- graph_from_data_frame(temp,directed = F)
  v_shape <- ifelse( names(V(g)) %in% unique(temp[,1]) , "square" , "circle" )
  v_color <- ifelse( names(V(g)) %in% unique(temp[,1]) , "honeydew" , "cornsilk" )
  v_size <- ifelse( names(V(g)) %in% unique(temp[,1]) , square , cricle )

  return(list(g=g, v_shape=v_shape, v_color=v_color, v_size=v_size))
  }

# Test parameters
# i <- "TF_PPI_Tissue"
# i <- "TF_TRRUST_enrichment"
# gr_name <- "Tissue"
# comp <- sheets

# Save all combination plots in one file
save_pdf.fun <- function(i, gr_name, comp){
  temp_i_name <- grep(i,list.files(paste0(out),full.names=F),value=T)
  p <- map(comp, ~read_xlsx(paste0(out, temp_i_name), sheet = .x))

  p_split <- p %>%
    imap(., ~.x %>%
          mutate(Regulation = paste0(.y, "_", .$Regulation)) %>%
          split(., .$Regulation)
          ) %>% compact() %>% flatten()

  obj_igraph <- imap(p_split, ~i_graph(.x, n, square, cricle))

  par(mar = c(0, 0, .7, 0))  #bottom, left, top, right
  pdf(file = paste0(out, "Plots/", i, "_network_",".pdf", sep=""))
    p <- imap(obj_igraph,
         ~with(.x, 
                 plot.igraph(g, label=V(g), page = T,
                 main = paste0(str_replace(.y, "_", " ")," regulated genes"),
                 layout=layout_with_kk(g), vertex.label.color="black" ,
                 vertex.size=v_size, vertex.label.cex=txt, vertex.label.font=2, 
                 vertex.shape=v_shape, vertex.color=v_color, marigin = c(-1,-4,0,0),
                 vertex.frame.color="grey60", curved=F)

    ))
    dev.off()
}

# if(isFALSE(dir.exists(paste0(out,"Plots")))){dir.create(paste0("../",result.dir),recursive = TRUE)}
# save_pdf.fun("TF_PPI_Luminal", "Luminal", comp = sheets[[1]])
# save_pdf.fun("TF_PPI_Tissue", "Tissue", comp = sheets[[2]])

@@
